Robustness of MinD oscillation in \emph{Escherichia coli} with diverse cell shapes

### 2014-09-19 paper/paper.tex
minor edits to Intro Tested everything on quipu.

Previous studies have shown experimentally that the MinD protein
system is capable of exhibiting oscillations in round
shapes~\cite{fange2006noise} as well as in connected three pronged
tube shapes~\cite{varma2008min}, in which the oscillations seem to
seek out the extreme poles in the cell~\cite{corbin2002exploring,
  juarez2010changes}.  One conclusion of these studies has been that
the MinD system robustly forms bipolar oscillations regardless of
variations in cellular shape.

However, Mannik \emph{et al.} have recently shown that there are
limitations to this robust capability to
oscillate~\cite{mannik2010bacteria, mannik2009bacterial}. They have
experimentally forced E. Coli cells into silicon microfabricated
channels of $0.25\micron$ thickness. Upon entering the channels, the
cells undergo a mechanical deformation in which they both widen and
lengthen within the plane of the channel.  This deformation results in
very wide (they can reach widths of over $5\micron$) flattened cells
that when viewed from the top down show irregular and asymmetric
shapes.  While these cells are still able to divide into surprisingly
equal volumes, the MinD oscillations in these cells is irregular,
both temporally and spatially. Seen with flourescent microscopy (after
using common genetic techniques to allow the cells to flouresce), the
MinD seem to maximize in multiple locations within the cell, in a
seemingly random sequence. This allows for an opportunity to test
previously studied MinD simulation models against more extreme
experimental cases than have been seen thus far.

A number of models of the MinD protein system have been developed,
which accurately describe the dynamics using simulations of protein
reaction and diffusion.
%
Early models involved free proteins that affect each others' rates of
diffusion and membrane attachment, but do not combine into compound
states~\cite{meinhardt2001pattern}.  In 2003 Huang improved upon this
work with a simple and very successful simulation model based on
MinD-MinE combination, ATPase hydrolysis, and MinD membrane attachment
that exhibits accurate MinD oscillations in cylindrical
cells~\cite{huang2003dynamic}. In this model cytoplasmic MinD is
recruited to the membrane by MinD is already clustered there
(following observed non-linear attachement of minD on the cell
membrane), and is stationary once attached.  A number of studies have
used an approach similar in that they do not rely on the ability of
MinD to move along the walls and
cluster~\cite{kruse2007experimentalist, meinhardt2001pattern,
  drew2005polymerization, fange2006noise, kerr2006division}, while
studies have been made as well of models which rely on MinD mobility
and attraction on the cell membrane~\cite{kruse2002dynamic,
  howard2005cellular}.  Experimental studies of the Min system's
association with the cell membrane have also been
made~\fixme{finding what results? We don't want to just list what has
  been done, but rather to summarize the state of
  knowledge.}~\cite{hsieh2010direct, mileykovskaya2003effects}.

>>>>

Previous studies have shown experimentally that the MinD protein
system is capable of exhibiting oscillations in round
shapes~\cite{fange2006noise} as well as in connected three pronged
tube shapes~\cite{varma2008min}, in which the oscillations seem to
seek out the extreme poles in the
cell~\cite{juarez2010changes,corbin2002exploring}.  One conclusion of
these studies has been that the MinD system robustly forms bipolar
oscillations regardless of variations in cellular shape.

However, Mannik \emph{et al.} have recently shown that there are
limitations to this robust capability to
oscillate~\cite{mannik2010bacteria, mannik2009bacterial}. They have
experimentally forced E. Coli cells into silicon microfabricated
channels of $0.25\micron$ thickness. Upon entering the channels, the
cells undergo a mechanical deformation in which they both widen and
lengthen within the plane of the channel.  This deformation results in
very wide (they can reach widths of over $5\micron$) flattened cells
that when viewed from the top down have irregular and asymmetric
shapes.  While these cells are still able to divide into surprisingly
equal volumes, the MinD oscillations in these cells are irregular,
both temporally and spatially. Seen with flourescent microscopy (after
using common genetic techniques to allow the cells to flouresce), the
MinD seem to maximize in multiple locations within the cell, in a
seemingly random sequence. This allows for an opportunity to test
previously studied MinD simulation models against more extreme
experimental cases than have been seen thus far.

A number of models of the MinD protein system have been developed,
which accurately describe the dynamics using simulations of protein
reaction and diffusion.
%
Early models involved free proteins that affect each others' rates of
diffusion and membrane attachment, but do not combine into compound
states~\cite{meinhardt2001pattern}.  In 2003 Huang improved upon this
work with a simple and very successful simulation model based on
MinD-MinE combination, ATPase hydrolysis, and MinD membrane attachment
that exhibits accurate MinD oscillations in cylindrical
cells~\cite{huang2003dynamic}. In this model cytoplasmic MinD is
recruited to the membrane by MinD that is already clustered there
(following observed non-linear attachement of minD on the cell
membrane~\cite{hu2002dynamic,shih2002division}), and is stationary
once attached.  A number of studies have used an approach similar in
that they do not rely on the ability of MinD to move along the walls
and cluster~\cite{kruse2007experimentalist, meinhardt2001pattern,
  drew2005polymerization, fange2006noise, kerr2006division}, while
studies have been made as well of models which rely on MinD mobility
and attraction on the cell membrane~\cite{kruse2002dynamic,
  howard2005cellular}.
